Fail Faster

### 2025-03-19 paper/paper.tex
more work on other langs

\begin{itemize}
  \item \textbf{Racket} is the next target that we intend to test the \name{} methodology on. The entire Racket philosophy is intertwined with
  using macros to build small eDSLs like the ones we use to write generators, and so it is a natural fit.
  We have not yet tested the technique in Racket for the simple reason that the authorship team of this paper does not posess
  much Racket expertise \jwc{can we say this without unblinding?}.

  \item \textbf{Haskell} birthed PBT, and so it is natural to ask why we have not yet applied \name{} to it. The basic answer is that
  because most tasks in Haskell are accomplished with monads, GHC is designed specially to eliminate the run-time overhead of monadic
  abstractions! Indeed, because Haskell is pure, GHC can aggressively inline and beta-reduce
  nearly anything as part of its normal optimiation steps.
  For complex enough programs however, the heuristics GHC uses can degrade, leading to performance overheads for monadic code.
  For this reason, Haskell does have a multi-stage programming system called Template Haskell \cn{}, which is sometimes 
  to build eDSLs in contexts where one does not want to rely on the optimizer's heuristics.
  \item \textbf{Rust:} asdf
  \item \textbf{Python:} asdf
\end{itemize}

\jwc{
  \begin{itemize}
    \item Racket is the only ``real'' strict functional language we haven't tested this in. It'd definitely work.
    \item Haskell GHC does a lot of these optimizations already, since the code is pure. Since QC generators are relatively small programs,
    GHC has little trouble specializing them. Of course, this is not guaranteed. A version of this idea can easily be ported to the original QC with template haskell, to guarantee
    the highest-performance generators. But the benefit would be lower since GHC is smarter. Worth doing, though.
    \item Rust: generator libraries are not monadic, presumably to avoid exactly these issues.
    \item Python. Maybe could do it?? Hypothesis is the most used PBT library out there so this could be high impact.
  \end{itemize}
}

@@

\subsection{Staging eDSLs}
\jwc{A generator is a parser of
randomness~\cite{goldsteinParsingRandomness2022}, and people have implemented
lots of staged parser libraries!}
\jwc{cite Lionel Parreaux  work, he's on PC}

>>>>

\begin{itemize}
  \item \textbf{Racket} is the next target that we intend to test the \name{} methodology on.
  The entire Racket philosophy is intertwined with
  using macros to build small eDSLs like the ones we use to write generators, and so it is a natural fit. Indeed,
  there are at least two \cn{} PBT libraries in Racket.
  \jwc{The Racket macros literature, while extensive, is not about staging!}
  We have not yet tested the technique in Racket for the simple reason that the authorship team of this paper does not posess
  much Racket expertise \jwc{can we say this without unblinding?}.

  \item \textbf{Haskell} birthed PBT, and so it is natural to ask why we have not yet applied \name{} to it. The basic answer is that
  the Haskell compiler (GHC) is designed specially to eliminate the run-time overhead of monadic
  abstractions! Indeed, because Haskell is pure, GHC can aggressively inline and beta-reduce
  nearly anything as part of its normal optimiation steps.
  This means that in many cases, the impact of a staged PBT library (and indeed, all staged monadic DSLs) negligible.
  For complex enough programs however, the heuristics GHC uses can degrade, leading to performance overheads for monadic code.
  For this reason, Haskell does have a multi-stage programming system called Template Haskell \cn{}, which is sometimes 
  to build eDSLs in contexts where one does not want to rely on the optimizer's heuristics.
  In short: it is possible to port \name{} to Haskell QuickCheck, but the benefits would be much less pronounced.

  \item \textbf{Rust} shares some similarities with functional languages, and
  indeed, it is host to a number of property-based testing libraries \cn{}. However Rust does not
  structurally encourage monadic eDSLs --- it does not have a special monadic syntax --- and so
  PBT libraries in Rust ask programmers to define generators directly as \texttt{seed -> value} functions.
  For this reason, \name{} does not seem directly applicable to any of the PBT libraries in Rust.

  \item \textbf{Python} is home to Hypothesis, the most widely used PBT library in any language.

\end{itemize}

@@

\subsection{Staging eDSLs}
\jwc{A generator is a parser of
randomness~\cite{goldsteinParsingRandomness2022}, and people have implemented
lots of staged parser libraries!}
\jwc{cite Lionel Parreaux  work, he's on PC}
\jwc{Macros in Racket}
